\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-12-11 chapters/bci.tex
Fix testprint issues

\Acp{bci} started as a speculative concept in the 1970's~\cite{Vidal1973}.
Now, they have gradually evolved from methods  developed in the the 1990's to
establish basic communication from minimal brain activity
(yes/no, left/right, \ldots)~\cite{Wolpaw2002} into sophisticated interaction
schemes that restore or enhance many capabilities by bypassing defective parts
of the nervous or muscle system.
By now, classic \acp{bci} methods are starting to mature outside the lab setting and can
be used as assistive communication technology by individuals with \acl{sspi}
~\cite{Wolpaw2018,Milekovic2018}.
Impressive cutting-edge experimental examples include a brain-spine interface allowing a
paraplegic individual to walk again~\cite{Lorach2023}, a \ac{bci} translating internal
speech to a facial avatar speaking the imagined text~\cite{Metzger2023}, and fast
communication through decoding imagined handwritten symbols~\cite{Willett2021}.
Recently, the idea has also gained popular traction through Elon Musk's
NeuraLink~\cite{Musk2019} initiative, which imagines a multipurpose \ac{bci}.

>>>>

\Acp{bci} started as a speculative concept in the 1970's~\cite{Vidal1973}.
Now, they have gradually evolved from methods  developed in the 1990's to
establish basic communication from minimal brain activity
(yes/no, left/right, \ldots)~\cite{Wolpaw2002} into sophisticated interaction
schemes that restore or enhance many capabilities by bypassing defective parts
of the nervous or muscle system.
By now, classic \acp{bci} methods are starting to mature outside the lab setting and can
be used as assistive communication technology by individuals with \acl{sspi}
~\cite{Wolpaw2018,Milekovic2018}.
Impressive cutting-edge experimental examples include a brain-spine interface allowing a
paraplegic individual to walk again~\cite{Lorach2023}, a \ac{bci} translating internal
speech to a facial avatar speaking the imagined text~\cite{Metzger2023}, and fast
communication through decoding imagined handwritten symbols~\cite{Willett2021}.
Recently, the idea has also gained popular traction through Elon Musk's
NeuraLink~\cite{Musk2019} initiative, which imagines a multipurpose \ac{bci}.

### 2024-12-11 chapters/gaze_independence.tex
Fix testprint issues

We believe it is generally not opportune to delve too deeply into the underlying
ophthalmological and neurological mechanisms for each of the specific
conditions, or even the exact symptoms.
In clinical reality, every \ac{bci} user has different symptoms, which lead to their own
unique set of preserved capabilities and visual skills.
From a solution-oriented \ac{bci} engineering point of view, the etiology of the symptoms can be abstracted away.
Instead of categorizing users by their etiology, we will refer to those who might benefit from
\ac{bci} assistive communication technology as having \ac{sspi}, in line with
the terminology used by~\textcite{FriedOken2020}.
If they have eye motor impairment that affects their use of visual \ac{bci} or
eye-tracking solutions, we use the term \ac{sspgi} (see \cref{fig:gaze-independence/venn}).

>>>>

We believe it is generally not opportune to delve too deeply into the underlying
ophthalmological and neurological mechanisms for each of the specific
conditions, or even the exact symptoms.
In clinical reality, every \ac{bci} user has different symptoms, which lead to their own
unique set of preserved capabilities and visual skills.
From a solution-oriented \ac{bci} engineering point of view, the etiology of the symptoms can be abstracted away.
Instead of categorizing users by their etiology, we will refer to those who might benefit from
\ac{bci} assistive communication technology as having \acf{sspi}, in line with
the terminology used by~\textcite{FriedOken2020}.
If they have eye motor impairment that affects their use of visual \ac{bci} or
eye-tracking solutions, we use the term \acf{sspgi} (see \cref{fig:gaze-independence/venn}).

### 2024-12-11 chapters/preface.tex
Fix testprint issues

Since my thesis was a joint PhD between two labs, one at KU Leuven in
Belgium and one at the University of Lille in France, I got to
experience working at two different research institutions
This formed an invaluable enrichment of my career until now.
It also means I've had the amazing luck of having double the amount of wonderful
colleagues.

@@

In Leuven, I want to begin with a special thanks to those of you who
guided me at the start of my PhD: Benjamin, for introducing me to
the wonderful world of ERPs and linear decoding, and Axel and Arno for
powering through COVID with me. As for the rest of you lot -- Aline,
Anahit, Aurélie, Barbara, Bob, Elvira, Eva, Li Ang, Liuyin, Mani, Mansoureh,
Qiang, Tjaša, Valentina, and Yide -- you've made my days brighter in many ways,
and I could always count on you, be it for advice, support or help.
I also wish to thank the students and interns I was fortunate enough to
supervise for their dedication and great work, as well as the students at NeuroTech Leuven.
You never failed to impress me with creative ideas and excellent
execution. A final shoutout to everyone in and around the
Neurophysiology lab: Alessia, Astrid, Christophe, Inez, Jonathan,
Kenneth, Lara, Noa, Saba, Sara, Sofie, Stijn, Wouter, and all others.

>>>>

Since my thesis was a joint PhD between two labs, one at KU Leuven in
Belgium and one at the University of Lille in France, I got to
experience working at two different research institutions.
This formed an invaluable enrichment of my career until now.
It also means I've had the amazing luck of having double the amount of wonderful
colleagues.

@@

In Leuven, I want to begin with a special thanks to those of you who
guided me at the start of my PhD: Benjamin, for introducing me to
the wonderful world of ERPs and linear decoding, and Arno and Axel for
powering through COVID with me. As for the rest of you lot -- Aline,
Anahit, Aurélie, Barbara, Bob, Elvira, Eva, Li Ang, Liuyin, Mani, Mansoureh,
Qiang, Tjaša, Valentina, and Yide -- you've made my days brighter in many ways,
and I could always count on you, be it for advice, support or help.
I also wish to thank the students and interns I was fortunate enough to
supervise for their dedication and great work, as well as the students at NeuroTech Leuven.
You never failed to impress me with creative ideas and excellent
execution. A final shoutout to everyone in and around the
Neurophysiology lab: Alessia, Astrid, Christophe, Inez, Jonathan,
Kenneth, Lara, Noa, Saba, Sara, Sofie, Stijn, Wouter, and all others.
